---
title: "5 Data Transformation Exercises"
author: "RH"
date: "2023-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

options(dplyr.summarise.inform = FALSE)

# install.packages("tidyverse")
# install.packages(c("gapminder", "Lahman", "nycflights13", "palmerpenguins", "wakefield"))

library(nycflights13)
library(tidyverse)
```

## 5.2.4 Exercises

# 1.
```{r}
# 1. Had an arrival delay of two or more hours
filter(flights, arr_delay >= 120)

# 2. Flew to Houston (IAH or HOU)
filter(flights, dest %in% c("IAH","HOU"))

# 3. Were operated by United, American, or Delta
# filter(filters, )

# 4. Departed in summer (July, August, and September)
filter(flights, month %in% c(7,8,9))

# 5. Arrived more than two hours late, but didn’t leave late
filter(flights, dep_delay == 0 & arr_delay >= 120)

# 6. Were delayed by at least an hour, but made up over 30 minutes in flight
filter(flights, dep_delay >= 60, arr_delay <= 30)

# 7. Departed between midnight and 6am (inclusive)
filter(flights, (hour >= 0 & hour <= 5 & minute >= 0 & minute <= 59) | (hour <= 6 & minute <= 0) )

```

# 2. 
```{r}
filter(flights, between(hour, 0, 6))

```

# 3.
```{r}
filter(flights, is.na(dep_time))

#8.255

```

# 4.
```{r}
# ???

```

## 5.3.1 Exercises

# 1.
```{r}
# How could you use arrange() to sort all missing values to the start? (Hint: use is.na())
arrange(flights, desc(is.na(flights)))

```

# 2.
```{r}
# Sort flights to find the most delayed flights. Find the flights that left earliest.
arrange(flights, desc(dep_delay), desc(arr_delay))
arrange(flights, dep_delay)

```
# 3
```{r}
# Sort flights to find the fastest (highest speed) flights.
arrange(flights, desc(distance / air_time))


```

# 4
```{r}
# Which flights travelled the farthest? Which travelled the shortest?
arrange(flights, desc(distance))

```

## 5.4.1 Exercises
# 1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)

vars <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
select(flights, any_of(vars))

```

# 2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, year, year)

```

# 3. What does the any_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

# flights %>% select(any_of(vars))
select(flights, any_of(vars))
```

# 4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
# Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
select(flights, contains("TIME"))

```

## 5.5.2 Exercises

# 1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
# select(flights, dep_time, sched_dep_time)

select(flights, contains("dep_time"))

transmute(flights,
          dep_time,
          dep_time_minutes = (dep_time %/% 100 * 60) + dep_time %% 100,
          sched_dep_time,
          sched_dep_time_minutes = (sched_dep_time %/% 100 * 60) + sched_dep_time %% 100
          )

```


# 2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}
# ?flights
transmute(flights,
          air_time,
          arr_time,
          dep_time,
          arr_time_min = (arr_time %/% 100 * 60) + arr_time %% 100,
          dep_time_min = (dep_time %/% 100 * 60) + dep_time %% 100,
          arr_dep_diff = arr_time_min - dep_time_min
          )

```


# 3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
```{r}
transmute(flights,
          dep_time,
          sched_dep_time,
          dep_delay,
          dep_time_min = (dep_time %/% 100 * 60) + dep_time %% 100,
          sched_dep_time_min = (sched_dep_time %/% 100 * 60) + sched_dep_time %% 100,
          dep_shed_dep_diff = dep_time_min - sched_dep_time_min
          )

```


# 4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

# 5. What does 1:3 + 1:10 return? Why?

# 6. What trigonometric functions does R provide?
```{r}

```

