---
title: "5 Exploratory Data Anlaysis"
author: "RH"
date: "2023-01-27"
output: html_notebook
---

```{r}
knitr::opts_chunk$set(echo = FALSE)

options(dplyr.summarise.inform = FALSE)

# install.packages("tidyverse")
# install.packages(c("gapminder", "Lahman", "nycflights13", "palmerpenguins", "wakefield"))
library(tidyverse)
```

##Exercises

#7.3.4 Exercises
1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
# ?diamonds

unusual <- diamonds %>% 
  filter(y < 3 | y > 20) %>% 
  select(price, x, y, z) %>%
  arrange(y)
unusual

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = x))

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = y))

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = z))

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = x), color = "blue") +
  geom_bar(mapping = aes(x = y), color = "red") +
  geom_bar(mapping = aes(x = z), color = "green")
```


2. Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)
```{r}
# ?geom_freqpoly

ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 5000)
```


3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?
```{r}
# ?diamonds

diamonds
filter(diamonds, carat == "0.99" )
filter(diamonds, carat == "1" )
```

4. Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?

```{r}
p <- ggplot(diamonds, aes(x = y)) +
  geom_histogram(binwidth = 0.5) 
p

p + coord_cartesian(xlim = c(0,50))
p + coord_cartesian(ylim = c(0,50))
```

#7.4.1 Exercises
1. What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?

2. What does na.rm = TRUE do in mean() and sum()?